The brightness and spatial distributions of terrestrial radio sources

### 2013-02-28 articles/Article-RFI-distribution/article-rfi-distribution.tex
Fixing Matthews comments

\section{Introduction}
Radio astronomy concerns itself with the observation of radiation from celestial sources at radio wavelengths. However, astronomical radio observations can be affected by radio-frequency interference (RFI), which makes it difficult to calibrate the instrument and achieve high sensitivities \citep{impact-of-warc79,interference-and-radioastronomy-1991,interference-model-lemmon,rfi-mitigation-overview-fridman-baan}. Many techniques have been designed to mitigate this effect, such as detection and flagging of data \citep{chi-square-time-blanking-weber, multichannel-rfi-mitigation, exoplanet-detection-with-rfi, wsrt-rfims, pulse-blanking, effelsberg-rfi-mitigation, post-correlation-rfi-classification}, adaptive cancellation techniques \citep{adaptive-cancellation,post-correlation-reference-signal} and spatial filtering \citep{multichannel-rfi-mitigation, ellingson-spatial-nulling-2002, hampson-spatial-nulling-2002, boonstra-dissertation, spatial-filtering-parkes-multibeam-for-pulses, post-correlation-filtering}. In most cases, and with the help of careful spectrum allocation management and radio-quiet zones, these methods excise enough of the interference to allow calibration and imaging of the data.

Typical radio observations record a few hours of data, and the results are integrated. In these cases, excising only the interference that is apparent and thus above the noise often suffices, i.e., the observation can still reach the thermal noise limit of the instrument. A new challenge arises, however, when one desires much deeper observations, and hundreds of hours of observations need to be integrated. In such a case, weak interference caused by stationary sources might not manifest itself above the noise in individual observations, but might be coherently present in the data. Subsequently, when averaging these data, the interference might become apparent and occlude the signal of interest. This is very relevant for the 21-cm Epoch of Reionisation (EoR) experiments, because they involve long integration times. Several such experiments are underway, to either measure the angular power spectrum \citep{gmrt-eor-2011-paciga,de-bruyn-eor-ursi-2011,eor-paper-2011-jacobs,eor-mwa-2012-williams} or the global signal \citep{edges}. Ground-based Cosmic Microwave Background (CMB) experiments are another class of experiments involving long integration times (e.g., \citealt{ska-cmb-subrahmanyan}). For these experiments, it is important to know the possible effect of low-level interferences on the data, as these might overshadow or alter the signal of interest. In this article, we will connect new insights about RFI to the angular Epoch of Reionisation experiment that is using the Low-Frequency Array (LOFAR) (\citealt{de-bruyn-eor-ursi-2011}, van Haarlem et al. 2013, \textit{submitted}).
% TODO correct reference if van Haarlem et al is published

>>>>

\section{Introduction}
Radio astronomy concerns itself with the observation of radiation from celestial sources at radio wavelengths. However, astronomical radio observations can be affected by radio-frequency interference (RFI), which might make it difficult to calibrate the instrument and achieve high sensitivities \citep{impact-of-warc79,interference-and-radioastronomy-1991,interference-model-lemmon,rfi-mitigation-overview-fridman-baan}. Many techniques have been designed to mitigate this effect, such as detection and flagging of data \citep{chi-square-time-blanking-weber, multichannel-rfi-mitigation, exoplanet-detection-with-rfi, wsrt-rfims, pulse-blanking, effelsberg-rfi-mitigation, post-correlation-rfi-classification}, adaptive cancellation techniques \citep{adaptive-cancellation,post-correlation-reference-signal} and spatial filtering \citep{multichannel-rfi-mitigation, ellingson-spatial-nulling-2002, hampson-spatial-nulling-2002, boonstra-dissertation, spatial-filtering-parkes-multibeam-for-pulses, post-correlation-filtering}. In most cases, and with the help of careful spectrum allocation management and radio-quiet zones, these methods excise enough of the interference to allow calibration and imaging of the data.

Typical radio observations record a few hours of data, and the results are integrated. In these cases, excising only the interference that is apparent and thus above the noise often suffices, i.e., the observation can still reach the thermal noise limit of the instrument. A new challenge arises, however, when one desires much deeper observations, and hundreds of hours of observations need to be integrated. In such a case, weak interference caused by stationary sources might not manifest itself above the noise in individual observations, but might be coherently present in the data. Subsequently, when averaging these data, the interference might become apparent and occlude the signal of interest. This is very relevant for the 21-cm Epoch of Reionisation (EoR) experiments, because they involve long integration times. Several such experiments are underway, to either measure the angular power spectrum \citep{gmrt-eor-2011-paciga,de-bruyn-eor-ursi-2011,eor-paper-2011-jacobs,eor-mwa-2012-williams} or the global signal \citep{edges}. Ground-based Cosmic Microwave Background (CMB) experiments are another class of experiments involving long integration times (e.g., \citealt{ska-cmb-subrahmanyan}). For these experiments, it is important to know the possible effect of low-level interference on the data, as these might overshadow or alter the signal of interest. In this article, we will connect new insights about RFI to the angular Epoch of Reionisation experiment that is using the Low-Frequency Array (LOFAR) (\citealt{de-bruyn-eor-ursi-2011}, van Haarlem et al. 2013, \textit{submitted}).
% TODO correct reference if van Haarlem et al is published
